Identity of fish fry from the “Tetí” fishery in Eastern Cuba

### 2021-09-07 analysis/paper/paper.Rmd
updatingMS&Ref

---
title: 'Identity of fish fry from the “Tetí” fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
  - Prosanta Chakrabarty:
      email: prosanta@lsu.edu
      institute: [UofO]
      correspondence: true
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '../templates/template.docx'
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '../templates/cell-structure-and-function.csl'
abstract: |
  Fisheries of juvenile and larval fishes (i.e., fish-fry fisheries) often exploit the migratory behavior of diadromous species on their return to rivers. Although these fisheries are common worldwide, fish-fry fisheries are subject to insufficient management and the identity of the target species is frequently unknown. The artisanal fish-fry fishery in Cuba is known as “Tetí”, and is restricted to the eastern region of the country. Fishing “Tetí” has been practiced for hundreds of years, but an increase in tourism has elevated demand in recent years and may pose a threat to the conservation of this natural resource. It remains unclear if “Tetí” refers to one or multiple species. The present study aims to identify the fish fry that make up the “Tetí” fishery in Eastern Cuba with genetic data. We first used a standard nucleotide BLAST to identify the sequences. Then, we recovered the phylogenetic position of these samples based on a maximum likelihood phylogenetic analysis that also used available sequences from GenBank. Here, we provide evidence that Busck´s Stone-Biting Goby, Sicydium buscki, is a component of the “Tetí” fishery in Cuba. However, whether this fishery includes one or more species remains a mystery until additional samples can be analyzed. However, with this species identification the first efforts towards effective management can be taken to manage this fishery. Collecting additional scientific and fishing data from the “Tetí” fishery and including them in local management plans will be an important strategy for avoiding overfishing this resource.  
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; DNA barcoding; Gobiidae

@@

PCR products were purified using ExoSAP-IT™ PCR Product Cleanup Reagent (Affymetrix Inc./ USB, Cleveland, OH, USA), and then sequenced on an ABI 3100 automated sequencer (Applied Biosystems). Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious] and submitted to GenBank (accession numbers: MZ848370– MZ848378). We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhangetal2000), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.

@@

Although other species have been mentioned as potentially being part of the “Tetí” fishery, this is the first report of any fish species being identified based on molecular data. Other countries where fish-fry fisheries also take place have reported more than one species of goby (Gobiidae) and/or sleeper making up the fishery (Eleotridae) [@castellanos2011goby; @asis2013first; @engman2019puerto]. Thus, additional surveys including more samples from different localities are needed to uncover the diversity of gobies in Eastern Cuba and the fish diversity comprising this particular fishery. An increase in tourism in Cuba, possibly leading to a higher demand for this natural resource, is a relatively new threat faced by the native fishes. They are also threatened by the introduction of exotic species, habitat fragmentation, and water pollution [@monitoreoCuba]. All this makes the present study a valuable tool providing basic information about the fish fauna of the region. More thorough investigations and identification of species comprising the “Tetí” fishery will allow us to better work toward their conservation and propose effective management actions.

>>>>

---
title: 'Identity of fish fry from the “Tetí” fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
  - Prosanta Chakrabarty:
      email: prosanta@lsu.edu
      institute: [UofO]
      correspondence: true
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '../templates/template.docx'
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '../templates/cell-structure-and-function.csl'
abstract: |
  Fisheries of juvenile and larval fishes (i.e., fish-fry fisheries) often exploit the migratory behavior of diadromous species on their return to rivers. Although these fisheries are common worldwide, fish-fry fisheries are subject to insufficient management and the identity of the target species is frequently unknown. The artisanal fish-fry fishery in Cuba is known as “Tetí”, and is restricted to the eastern region of the country. Fishing “Tetí” has been practiced for hundreds of years, but an increase in tourism has elevated demand in recent years and may pose a threat to the conservation of this natural resource. It remains unclear if “Tetí” refers to one or multiple species. The present study aims to identify the fish fry that make up the “Tetí” fishery in Eastern Cuba with genetic data. We first used a standard nucleotide BLAST to identify the sequences. Then, we recovered the phylogenetic position of these samples based on a maximum likelihood phylogenetic analysis that also used available sequences from GenBank. Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the “Tetí” fishery in Cuba. However, whether this fishery includes one or more species remains a mystery until additional samples can be analyzed. However, with this species identification the first efforts towards effective management can be taken to manage this fishery. Collecting additional scientific and fishing data from the “Tetí” fishery and including them in local management plans will be an important strategy for avoiding overfishing this resource.  
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; DNA barcoding; Gobiidae

@@

PCR products were purified using ExoSAP-IT™ PCR Product Cleanup Reagent (Affymetrix Inc./ USB, Cleveland, OH, USA), and then sequenced on an ABI 3100 automated sequencer (Applied Biosystems). Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious] and submitted to GenBank (accession numbers: MZ848370– MZ848378). We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhang2000greedy), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.

@@

Although other species have been mentioned as potentially being part of the “Tetí” fishery, this is the first report of any fish species being identified based on molecular data. Other countries where fish-fry fisheries also take place have reported more than one species of goby (Gobiidae) and/or sleeper making up the fishery (Eleotridae) [@castellanos2011goby; @asis2013first; @engman2019puerto]. Thus, additional surveys including more samples from different localities are needed to uncover the diversity of gobies in Eastern Cuba and the fish diversity comprising this particular fishery. An increase in tourism in Cuba, possibly leading to a higher demand for this natural resource, is a relatively new threat faced by the native fishes. They are also threatened by the introduction of exotic species, habitat fragmentation, and water pollution [@monitoreoCuba2017peces]. All this makes the present study a valuable tool providing basic information about the fish fauna of the region. More thorough investigations and identification of species comprising the “Tetí” fishery will allow us to better work toward their conservation and propose effective management actions.
